---
title: "XLSForm"
subtitle: "Tips, tricks, gotchas, and workarounds"  
author: "Ernest Guevarra"
date: '23 February 2021'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["right", "bottom"]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

## Libraries
library(ruODK)
library(dplyr)
library(knitr)
library(kableExtra)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  #base_color = "#509935",
  base_color = "#000000",
  base_font_size = "22px",
  title_slide_background_color = "#FFFFFF",
  title_slide_background_image = "images/ecohealth_title_background_4by3.png",
  title_slide_background_size = "contain",
  #title_slide_text_color = "#509935",
  title_slide_text_color = "#000000",
  background_image = "images/ecohealth_slide_background_4by3.png",
  header_font_google = google_font("Fira Sans"),
  text_font_google = google_font("Fira Sans Condensed"),
  code_font_google = google_font("Fira Mono"),
  code_font_size = "0.7rem",
  text_slide_number_font_size = "1em"
)
```


# Outline

* Background on XLSForm

* Guiding principle in form authoring

* XLSForm topics

    - Style recommendations

    - Optimising GPS in forms

    - Getting to know `select_multiple`
    
---

# Background

* form standard used to author Open Data Kit (ODK) forms

* uses a spreadsheet (Microsoft Excel, Google Sheets, or OpenDocument)

* simple to get started; sophisticated enough to allow complex authoring of forms

---

class: center, middle

# Guiding principle

## Author forms with data analysis in mind

---

class: center, middle

# XLSForm topics

[Link to XLSForm for this talk](https://docs.google.com/spreadsheets/d/1-u3ogfGSWbcb1m9yduWyXgXx9WTCAl56Mk8vMl9w_GY/edit?usp=sharing)

---

# Style recommendations

* Creating blank base form templates - see [this](https://docs.google.com/spreadsheets/d/1Hkogz5B92OnyyHt7C-bAXpB4md2jPXFVWecebGlQwGs/edit?usp=sharing)

* A fixed-width font is an ideal font to use - e.g., Courier, Menlo, Monaco, Roboto Mono

* Adding colour theme can be helpful to delineate sections of your form - helpful when visualising your form structure and relationthips of the different sections

---

# Optimising GPS in forms

* Early days of ODK, GPS capabilities were not as good

* Better now from ODK capture perspective but still mainly dependent on the GPS capability of the mobile device

* If collecting GPS, prime your device and prime ODK

* If accuracy is critical, set GPS with accuracy threshold

---

# Getting to know `select_multiple`

* Data structure associated with `select_multiple`

* _When to_ and _when not to_ use `select_multiple` - some recommendations

* Other interesting usage of `select_multiple`

